---
title: "Create Geographic Extent"
author: "steppe"
date: "2023-09-02"
output: html_document
---
 
```{r load libraries and import functions}
library(tidyverse)
library(sf)
source('functions.R')
```

The first step in generating a species list for the entirety of the Gunnison Basin is to create a spatial extent which represents the study area. Here we use the (Major River Basin Boundaries)[https://cdss.colorado.gov/gis-data/gis-data-by-category] from the Colorado Decision Support System on September 19, 2023.

```{r Identify drainage area}

watershed <- st_read('../data/raw/geodata/All_River_Basins/All_River_Basins.shp', quiet = T)  %>% 
  filter(HU6NAME == 'Gunnison')

```

These data were subset to the Gunnison Basin, buffered by 10km, and then a convex hull was rendered around the extent. 
The convex hull was used to simplify the shape for geospatial requests to subsequent databases, allowing for faster processing times. 
The hull was then separated into four parts, using 'st_make_grid' from the sf package, to allow for parcels sizes appropriate for query with the SORO herbarium consortia. 

```{r create grids for soro download, eval = F}

ws_buf <- st_buffer(watershed, 10000) %>% 
  st_convex_hull()

cells <- st_make_grid(ws_buf, n = c(2,2)) %>% 
  st_intersection(., ws_buf) %>% 
  st_transform(4326)

ggplot() +
  geom_sf(data = watershed) +
  geom_sf(data = cells)

```

These shapes were converted into well-known-text (WKT) format using the 'st_as_text' function. 
These four shapes were then manually queried into the Consortium of Southern Rockies herbaria by hand, because support for an API was undocumented. 
The data were downloaded as Symbiota native, without image records, occurrence traits attributes, or material samples as ISO-8859-1 for compliance with the R statistical language.

```{r create well known text for soro download, eval = F}
wkt <- st_as_text(cells)

coords <- lapply(wkt, lat_first)
coords <- unlist(coords)
write.table(coords, file = '../data/processed/wkt_boundaries.txt', col.names = F, row.names = F, quote = F)
```

In order to characterize species occurrences at finer resolutions the USGS National Hydrologic Database was intersected to the extent of the Gunnison Basin.

```{r subset data to fine resolution drainages, eval = F}

fine_watersheds <- st_read('../data/raw/geodata/WBD_14_HU2_Shape/Shape/WBDHU12.shp', quiet = T) 
watershed <- st_transform(watershed, st_crs(fine_watersheds))

fine_watersheds <- st_intersection(watershed, fine_watersheds) %>% 
  select(name)

ggplot() +
  geom_sf(data = fine_watersheds)

dir.create('../data/processed/fine_watersheds')
st_write(fine_watersheds, '../data/processed/fine_watersheds/fine_watersheds.shp')

rm(fine_watersheds, watershed)
```

